|

### 2022-11-07 manuscript.Rmd
finish intro update

In fact, training on accessibility has been increasingly becoming an important job-readiness aspect in computing and information sciences [@shinoharaWhoTeachesAccessibility2018; @kawasTeachingAccessibilityDesign2019]. For instance, some notable big tech companies, such as Yahoo!, Facebook, Microsoft, Google, and Adobe, have been leading a TeachAccess (teachaccess.org) Consortia to represent this increased demand for accessibility knowledge and to increase the number of computer and information science faculty teaching accessibility [-@shinoharaWhoTeachesAccessibility2018]. The Partnership on Employment & Accessible Technology (PEAT), in their Accessible Technology Skills Gap Report [-@peat], also highlights this trend--84% of industry correspondents that they work with say "it is important or very important for them to hire developers and designers with accessible technology skills." Growing interest in and importance of teaching accessibility can be also found in professional societies. For example, accessibility has been recently added to engineering design accreditation requirements by Accreditation Board for Engineering and Technology [ABET; abet.org] [@shinoharaWhoTeachesAccessibility2018].

Even though this statistic may be different specifically for data scientists, the data science as a highly interdisciplinary field across computing and information sciences also shares the rising demand for the accessibility preparedness. Furthermore, data science is broad with the intersection in careers as developers (e.g., data dashboard and package developers) and designers (e.g., data visualization experts). Therefore, the impact of teaching and learning accessibility is as far-reaching as its wide scope.

The PEAT also states that 60% of industry respondents "it was difficult or very difficult for their organization to find job candidates with accessibility skills."

Despite the increasing demand for accessibility knowledge in both industries and academia, accessibility from disabilities perspectives has not yet found its place in the data science course materials.
One main reason for this could be that accessibility is often covered in the computer sciences [@kawasTeachingAccessibilityDesign2019] or the digital arts [@barata2019inclusion], if at all covered.
However, data science is an interdisciplinary field that is taught in various disciplines, including computer science, statistics, business schools, political science, and biological sciences [@schwab2021data].
Thus, we believe that the resources and conversations around accessibility have not yet extended to the broader data science education community at the scale that they should.
For instance, many introductory data science courses [@yan2019first; @baumer2015data], textbooks [@adhikari2019computational; @wickham2016r], or curricula [@schwab2021data] either fail to mention accessibility at all or only focus on the accessing data.

In this manuscript, we argue that accessibility should be taught early in the data science curriculum so that regardless of whether learners major in data science or not, they can have foundational exposure to accessibility.
As data science educators who teach accessibility as part of our lower-division courses in two different institutions, we share specific examples that can be utilized by other data science instructors.
We teach R [@R-base] as the main language in our courses; thus, our examples are shaped around using R. However, instructors using a different set of language(s) or even teaching language-agnostic courses may still find the content applicable for their own courses.

@@

Although there are various ways to represent data patterns and models, visualization has been primarily taught in many data science courses for its efficiency [@kimAccessibleVisualizationDesign2021]. 
Such vision-dependent output may cause critical barriers against blind and visually impaired people and sometimes people with learning disabilities [@kimAccessibleVisualizationDesign2021; @marriottInclusiveDataVisualization2021; @leeReachingBroaderAudiences2020]. 
We argue that instructors need to teach multiple data representation methods so that all students can produce data products that are more accessible. 
To this end, we have considered multi-data representations and their relation to accessibility in our courses, and that's what we will demonstrate throughout this paper.

>>>>

In fact, training on accessibility has been increasingly becoming an important job-readiness aspect in computing and information sciences [@shinoharaWhoTeachesAccessibility2018; @kawasTeachingAccessibilityDesign2019]. For instance, some notable big tech companies, such as Yahoo!, Facebook, Microsoft, Google, and Adobe, have been leading a TeachAccess (teachaccess.org) Consortia to represent this increased demand for accessibility knowledge and to increase the number of computer and information science faculty teaching accessibility [-@shinoharaWhoTeachesAccessibility2018]. The Partnership on Employment & Accessible Technology (PEAT), in their Accessible Technology Skills Gap Report [-@peat], also highlights this trend--84% of industry correspondents that they work with say "it is important or very important for them to hire developers and designers with accessible technology skills." The growing interest in and importance of teaching accessibility can also be found in professional education societies. For example, accessibility has been recently added to engineering design accreditation requirements by Accreditation Board for Engineering and Technology [ABET; abet.org] [@shinoharaWhoTeachesAccessibility2018].

Data science consists of an iterative process of importing, tidying, transforming, visualizing, and modeling data [@wickham2016r]. While each stage has unique challenges, data visualization may pose fundamental barriers against people with permanent, situational, and temporary disabilities (e.g., visual, cognitive, and motor impairments, to name a few) in accessing information [@marriottInclusiveDataVisualization2021; @leeReachingBroaderAudiences2020]. For example, people with cognitive and learning disabilities often have difficulties understanding the abstractions and symbolic conventions used in visualizations, and people lacking fine motor coordination or having tremors may be unable to operate interactive visualizations, not to mention people who are blind and have low vision facing extra challenges in visualization access [@marriottInclusiveDataVisualization2021]. @marriottInclusiveDataVisualization2021, in their call to action paper, highlights the serious impact of inaccessible visualizations: "This is an equity issue with severe consequences, because a lack of access to visualizations and underlying data impacts health, educational and work opportunities, and lifestyle." <!-- Need to find a specific page number for this quote. -->

Despite the increasing demand for accessibility knowledge in both industries and academia and the critical impact of accessible visualizations, accessibility has not yet found its place in the data science course and teaching materials. One main reason for this could be that accessibility is often incorporated into human-computer interaction (HCI) and design courses [@shinoharaHowDesigningPeople2016; @shinoharaTeachingAccessibilityTechnology2017a], introductory web development [@koAccessComputingPromotesTeaching2016; @rosmaitaAccessibilityFirstNew2006], software engineering courses [@martin-escalonaIntroductionTopicAccessibility2013; @ludiIntroducingAccessibilityRequirements2007; @ludiAccessEveryoneIntroducing2002] or the digital arts [@barata2019inclusion], if at all covered. However, data science is an interdisciplinary field that is taught in various disciplines, including not only computer and information sciences but also statistics, business schools, political science, biological sciences, and more [@schwab2021data]. Thus, we posit that the resources and conversations around accessibility have not yet extended to the broader data science education community at the scale that they should.
For instance, many introductory data science courses [@yan2019first; @baumer2015data], textbooks [@adhikari2019computational; @wickham2016r], or curricula [@schwab2021data] either fail to mention accessibility at all or only focus on the accessing data.

In response to the need to develop accessible teaching materials within the data science discipline, this paper demonstrates and shares some teaching strategies that we have employed in our teaching introductory data science courses. In this manuscript, we argue that accessibility should be taught early in the data science curriculum so that regardless of whether learners major in data science or not, they can have foundational exposure to accessibility.

Some people may wonder why we need to teach accessibility as part of an introductory data science course as opposed to covering accessibility in a separate and general course module. According to the recent faculty survey on teaching accessibility [@shinoharaWhoTeachesAccessibility2018], the most critical barriers to teaching accessibility reported by 1,857 faculty from 318 institutions were "the absence of clear and *discipline-specific accessibility* [emphasis added] learning objectives and the lack of faculty knowledge about accessibility." <!-- Need to find the exact page number for this quote -->
This means faculty often have trouble making sense of how they can integrate accessibility into their specific teaching domains, which requires applied knowledge of accessibility beyond general concepts.

We share our experiences and strategies in the hope that our discipline-specific (i.e., data science and data visualization) accessibility teaching cases can be used as scaffolding guides for other data science instructors to build upon.

Another critique we may receive includes the timing of when to introduce accessibility concepts in data science courses. For example, some instructors may prefer to put teaching accessibility on the back burner so that they can save more time for their students to learn data science fundamentals first and adopt accessibility principles in an upper-level course later. While the decision is subject to each institution's and individual instructor's discussion, we believe the sooner, the better. This is because: (1) some students may take an introductory level only, which means they will lose a chance to get exposed to this demanding knowledge unless they elect the upper-level course covering accessibility; (2) this kind of separation may unintentionally contribute to misconceptions, such as accessibility is difficult and/or accessibility is optional, not essential. When properly incorporated, accessibility can be taught along with other data science fundamentals (e.g., wrangling, merging, tidying, visualizing) while not necessarily putting more cognitive burdens against students' current learning. Rather, students can take a chance to become creative design thinkers by making their data science work more meaningful and accessible through "empathy learning" [@putnamTeachingAccessibilityLearning2015]. Thus, we suggest teaching accessibility as part of an introductory data science course for students to practice coupling their data science knowledge with accessibility principles as early as possible.

As data science educators who teach accessibility as part of our lower-division courses in two different institutions, we share specific examples that can be utilized by other data science instructors.
We teach R [@R-base] as the main language in our courses; thus, our examples are shaped around using R. However, instructors using a different set of language(s) or even teaching language-agnostic courses may still find the content applicable to their own courses.

@@

Although there are various ways to represent data patterns and models, visualization has been primarily taught in many data science courses for its efficiency [@kimAccessibleVisualizationDesign2021]. 
Such vision-dependent output may cause critical barriers against not only blind and visually impaired people but also  people with cognitive and learning impairments and people having motor impairments [@kimAccessibleVisualizationDesign2021; @marriottInclusiveDataVisualization2021; @leeReachingBroaderAudiences2020]. 
We argue that instructors need to teach multiple data representation methods so that all students can produce data products that are more accessible. 
To this end, we have considered multi-data representations and their relation to accessibility in our courses, and that's what we will demonstrate throughout this paper.
